---
title: "Acquire"
---

## Overview

The aim of this script is to acquire the data from the ACTIV-ES project. The data is hosted on GitHub <https://github.com/francojc/activ-es>. The data is downloaded from the GitHub repository and saved in the `data/original` folder.

There are three versions of this data:

- `plain.zip`
  - Running test version.
- `tagged.zip`
  - POS tagged version ([Penn tagset](https://userpage.fu-berlin.de/~structeng/wiki/doku.php?id=corpora:tagset-penn)).
- `eagles.zip`
  - POS tagged version ([EAGLES tagset](https://www.sketchengine.eu/spanish-freeling-part-of-speech-tagset/) or the [Freeling repo](https://github.com/TALP-UPC/FreeLing/blob/master/doc/tagsets/tagset-es.html)).

Here are some considerations for our approach:

- One use of this data will be to demonstrate working with running, unstructured text. So we will use the `plain.zip` version of the data.

- Another potential use would be to explore the verbal paradigms of the verbs in the data. For this we would need the POS tagged version of the data. The `eagles.zip` version of the data is tagged with the EAGLES tagset (which is very robust, more so than the Penn tagset).

## Acquire

So the idea is to download the `plain.zip` version and the `eagles.zip` version of the data. We will save the `plain.zip` version and the `eagles.zip` version in the `data/original` folder. Each will have the own subfolders: `plain` and `eagles`.

Let's load the key packages for this task.

```{r}
#| label: acquire-load-packages

# Load packages
library(dplyr)        # for data manipulation
library(qtkit)        # for get_archive_data(), create_data_origin()
                      # functions
library(readr)        # for read_csv(
library(knitr)        # for kable()
library(fs)           # for dir_tree()
```

Now to download and write the data to the respective folders.

```{r}
#| label: acquire-download-data

# Set the URL for the datasets
plain_url <- "https://github.com/francojc/activ-es/raw/master/activ-es-v.02/corpus/plain.zip"

eagles_url <- "https://github.com/francojc/activ-es/raw/master/activ-es-v.02/corpus/eagles.zip"

# Download, unarchive, and write the data
get_archive_data(plain_url,
  target_dir = "../data/original/plain/",
  confirmed = TRUE
)
get_archive_data(eagles_url,
  target_dir = "../data/original/eagles/",
  confirmed = TRUE
)
```

The data for the `plain` and `eagles` versions of the data have been downloaded and saved in the `data/original` folder.

## Data origin

```{r}
#| label: create-data-origin
#| eval: false

# Create data origin (interactively)
create_data_origin(
  file_path = "../data/original/actives_do.csv"
)
```

View the data origin after creating and editing it.

```{r}
#| label: tbl-view-data-origin
#| tbl-cap: "Data origin for the ACTIV-ES project."
#| tbl-colwidths: [25, 75]
#| message: false

# View data origin
read_csv("../data/original/actives_do.csv") |>
  kable()
```

## Directory structure

```{r}
#| label: view-directory-structure

# View the directory structure
dir_tree("../data")
```
